---
title: "Data Import Script"
format: html
---

```{r}
library(rvest)
library(tibble)
library(stringr)
library(tidyverse)
library(httr2)
library(httr)
library(purrr)
library(dplyr)
```

## Scraping Country Names

This code was originally intended to scrape country names and food prices from Numbeo. Although it was successfully collected the list of countries, I wasn't able to scrape the detailed food price data because the website began blocking the request. I switched to using Numbeo's API to obtain market/restaurant price data instead. Fortunately, the country names I collected from the original scraping code were still useful later in the project, especially when mapping price across all countries.

```{r}
url <- "https://www.numbeo.com/food-prices/"

##scrape country names
country_names <-
  read_html(url)|>
  html_elements("table.related_links")|>
  html_elements("a")|>
  html_text2()

##scrape country URLs
country_links <-
  read_html(url)|>
  html_elements("table.related_links")|>
  html_elements("a")|>
  html_attr("href")|>
  url_absolute(base=url)
  
## remove "s" in url links
country_links <- str_replace(country_links, "^https", "http")

countries <- tibble(
  country = country_names,
  url = country_links
)

countries
```

## Scraping Food Price Data From Each Country

```{r}
get_food_prices <- function(country_url) {
  Sys.sleep(time = 30)
  page <- read_html(country_url)
  
  items <- page|>
    html_elements("tr.tr_standard, tr.tr_highlighted")|>
    html_elements("td:nth-child(1)")|>
    html_text2()
  
  prices <- page|>
    html_elements("tr.tr_standard, tr.tr_highlighted")|>
    html_elements("td:nth-child(2)")|>
    html_text2()|>
    str_remove("\\$")|>
    str_trim()
  
  tibble(
    item = items,
    price_usd = prices
  )
}

# food_data_all <- countries |>
#   mutate(food_data = map(url, get_food_prices, .progress = TRUE)) |>
#   unnest(food_data)

food_data <- map(countries$url, get_food_prices, .progress = TRUE)

food_data_all
```

Below is the code for retrieving Market/Restaurant prices using Numebeo API.

## Retrieve Market/Restaurant Price Data Using Numbeo API

```{r}
base_url <- "https://www.numbeo.com/api"

get_food_prices <- function(country) {
  Sys.sleep(1)

  response <- request(base_url) |>
  req_url_path_append("country_prices") |>
  req_url_query(api_key = "i987lirw7h13dn", country = country, currency = "USD") |>
  req_perform()
  food_prices_list <- resp_body_json(response)
  return(as_tibble(food_prices_list))
}

countries_with_prices <- map(countries$country, get_food_prices)
combined_food_prices <- bind_rows(countries_with_prices, .id = "country")
head(combined_food_prices)
```

## Scraping Income by Countries

```{r}
url_2 <- "https://worldpopulationreview.com/country-rankings/median-income-by-country"

income <- read_html(url_2) |>
  html_elements(".wpr-table")|>
  html_table(fill = TRUE)

income_table <- income[[1]]
str(income_table)
```

## Scraping Obesity Data

```{r}
url_3 <- "https://worldpopulationreview.com/country-rankings/obesity-rates-by-country#obesity-by-country"

obesity <- read_html(url_3)|>
  html_table()

obesity_table <- obesity[[1]]
str(obesity_table)
```
